---
title: "Homework 5"
author: "Khang Thai"
date: "2025-05-31"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1)

```{r}
library(RMariaDB)
library(DBI)
con <- dbConnect(RMariaDB::MariaDB(),
host = "relational.fel.cvut.cz",
port = 3306,
username = "guest",
password = "ctu-relational",
dbname = "SFScores"
)

```

### (a)

```{r}
dbGetQuery(con, "
  SELECT avg(avg_day_between)
  FROM (
    SELECT business_id,
      DATEDIFF(MAX(date), MIN(date)) / (COUNT(*) - 1) AS        avg_day_between
    FROM inspections
    WHERE type = 'Routine - Unscheduled'
    GROUP BY business_id
    HAVING COUNT(*) > 1
  ) AS avg_intervals;
")

```

### (b)

```{r}
monthly_score <- dbGetQuery(con, "
  SELECT MONTH(date) AS month, AVG(score) AS avg_score
  FROM inspections
  WHERE score IS NOT NULL
  GROUP BY month
")
monthly_score
cor(monthly_score$month, monthly_score$avg_score)
```

### (c)

```{r}
dbGetQuery(con, "
  WITH inspection_score AS (
    SELECT business_id, date, score,
    LAG(score) OVER (PARTITION BY business_id ORDER BY date) AS previous_score
    FROM inspections
    WHERE score IS NOT NULL
  ), businesses_with_drop AS (
    SELECT DISTINCT business_id
    FROM inspection_score
    WHERE score < previous_score
  ), nondecrease_or_uninspected AS (
    SELECT COUNT(*) AS count_nondecrease
    FROM businesses
    LEFT JOIN businesses_with_drop ON businesses.business_id = businesses_with_drop.business_id
    WHERE businesses_with_drop.business_id IS NULL
  ), total_business_count AS (
    SELECT COUNT(*) AS total_count
    FROM businesses
  )
  SELECT n.count_nondecrease, t.total_count,
  ROUND(n.count_nondecrease * 100.0 / t.total_count, 2) AS percentage_nondecrease
  FROM nondecrease_or_uninspected n, total_business_count t;
")


```

## Question 2)

### (a)

```{r}
data <- dbGetQuery(con, "
  WITH ranked_inspections AS (
    SELECT i.business_id, i.date, i.score, COUNT(score) AS n_violations, 
      SUM(CASE WHEN v.risk_category = 'Low Risk' THEN 1 ELSE 0 END) AS n_low_risk,
      SUM(CASE WHEN v.risk_category = 'Moderate Risk' THEN 1 ELSE 0 END) AS n_moderate_risk,
      SUM(CASE WHEN v.risk_category = 'High Risk' THEN 1 ELSE 0 END) AS n_high_risk,
      RANK() OVER (PARTITION BY i.business_id ORDER BY i.date DESC) as rn 
  FROM inspections i
  JOIN violations v ON i.business_id = v.business_id
  GROUP BY i.business_id, i.date, i.business_id, i.score
  )
  SELECT business_id, date, n_violations, n_low_risk, n_moderate_risk, n_high_risk
  FROM ranked_inspections
  WHERE rn = 1
  ORDER BY n_violations DESC
  LIMIT 10;
")
data

plot(data$n_violations, data$score, main = "Score vs. Violations", xlab = "Violations", ylab = "Score")

```

Those with a higher score tend to have a lower violation risk than those who have a lower score.

### (b)

```{r}
dbGetQuery(con, "
  WITH violation_categorized AS (
    SELECT 
      CASE
        WHEN LOWER(v.description) LIKE '%food%' THEN 'food'
        WHEN LOWER(v.description) LIKE '%plumbing%' THEN 'plumbing'
        WHEN LOWER(v.description) LIKE '%utensils%' THEN 'utensils'
        ELSE 'other'
      END AS category
    FROM violations v
    JOIN inspections i ON v.business_id = i.business_id
    WHERE i.score <= 80
  )
  SELECT category, 
    COUNT(*) AS count, 
    ROUND(100.0 * COUNT(*) / SUM(COUNT(*)) OVER (), 2) AS percentage
  FROM violation_categorized
  GROUP BY category;
")

```

### (c)

```{r}
dbGetQuery(con, "
  SELECT i.business_id, b.name, COUNT(*) AS total_inspection
  FROM inspections i
  JOIN businesses b ON i.business_id = b.business_id
  LEFT JOIN violations v ON i.business_id = v.business_id 
  WHERE v.business_id IS NULL
  GROUP BY i.business_id, b.name
  ORDER BY total_inspection DESC
  LIMIT 10;
")

```

### (d)

```{r}
dbGetQuery(con, "
  WITH inspection_rank AS (
    SELECT b.postal_code, i.score,
      RANK() OVER (PARTITION BY b.postal_code ORDER BY i.score) AS rn,
      COUNT(*) OVER (PARTITION BY b.postal_code) AS total_count
    FROM businesses b
    JOIN inspections i on b.business_id = i.business_id
  ),
  at_least_30 AS (
    SELECT *
    FROM inspection_rank
    HAVING total_count >= 30
  ),
  median_score AS (
    SELECT postal_code, score AS median_score
    FROM at_least_30
    WHERE rn = (total_count + 1) / 2
  )
  SELECT postal_code, median_score
  FROM median_score
  ORDER BY median_score DESC
  LIMIT 10;
  
")

```

## Question 3)

```{r}
library(reticulate)
virtualenv_install("stats167_venv", packages = c("pandas", "matplotlib", "seaborn"))
use_virtualenv("stats167_venv", required = TRUE)

```

### (a)

```{python}
import sqlite3
import pandas as pd

con = sqlite3.connect("rideshare.db")

query = """
  WITH hourly_count AS (
    SELECT sub_type,
      strftime('%Y-%m-%d', start_date) AS start_day,
      strftime('%H', start_date) AS start_hour,
      COUNT(*) AS n_hourly_trips
    FROM trips
    WHERE strftime('%Y', start_date) = '2012'
      AND strftime('%m', start_date) BETWEEN '05' AND '11'
    GROUP BY sub_type, start_day, start_hour
  ), avg_trips AS(
    SELECT sub_type, start_day, start_hour, n_hourly_trips,
      AVG(n_hourly_trips) OVER (
      PARTITION BY sub_type, start_hour
      ORDER BY start_day ROWS BETWEEN 27 PRECEDING 
      AND CURRENT ROW ) AS rolling_avg_28d_trips
  FROM hourly_count
  ), ranked AS (
    SELECT *,
      RANK() OVER (PARTITION BY sub_type 
      ORDER BY rolling_avg_28d_trips DESC) as rnk
    FROM avg_trips
  )
  SELECT sub_type, start_day, start_hour, n_hourly_trips,        rolling_avg_28d_trips
  FROM ranked
  WHERE rnk <= 5;
"""
df = pd.read_sql_query(query, con)
print(df)
df['start_day'] = pd.to_datetime(df['start_day'])
df['start_hour'] = df['start_hour'].astype(int)

```

```{python}
import seaborn as sns
import matplotlib.pyplot as plt

plt.figure(figsize=(16, 6))
sns.lineplot(data=df, x='start_day', y='rolling_avg_28d_trips', hue='sub_type')
plt.title('28-Day Rolling Average of Hourly Bike Trips (May–Nov 2012)')
plt.xlabel('Date')
plt.ylabel('Rolling Avg (Hourly Trips)')
plt.legend(title='User Type')
plt.tight_layout()
plt.show()

```

Registered users consistenly have a higher and more stable 28-day rolling average of bike trips compared to casual users. Casual users show more fluctuation.

### (b)

```{python}
con = sqlite3.connect("rideshare.db")
query = """
  WITH start_counts AS (
    SELECT 
      start_station AS station,
      strftime('%Y-%m', start_date) AS month,
      COUNT(*) n_start_trips
    FROM trips
    GROUP BY station, month
  ), end_counts AS (
    SELECT
      end_station AS station,
      strftime('%Y-%m', end_date) AS month,
      COUNT(*) AS n_end_trips
    FROM trips
    GROUP BY station, month
  ), month_stats AS (
    SELECT 
      s.station, 
      s.month, 
      s.n_start_trips, 
      e.n_end_trips
    FROM start_counts s
    LEFT JOIN end_counts e ON s.station = e.station 
      AND s.month = e.month
  ), avg_stats AS (
    SELECT station,
      AVG(n_start_trips) AS avg_monthly_start_trips,
      AVG(n_end_trips) AS avg_monthly_end_trips,
      CASE
        WHEN AVG(n_end_trips) = 0 THEN NULL
        ELSE CAST(AVG(n_start_trips) AS FLOAT) / 
        AVG(n_end_trips)
      END AS avg_start_end_ratio
    FROM month_stats
    GROUP BY station
  ) SELECT
      station,
      avg_monthly_start_trips,
      avg_monthly_end_trips,
      avg_start_end_ratio
    FROM avg_stats
    WHERE avg_start_end_ratio IS NOT NULL
    ORDER BY avg_start_end_ratio DESC
    LIMIT 6;

"""

df = pd.read_sql_query(query, con)
print(df)

```

## Question 4)

### (a)

NoSQL - Document database: Since the data is user-specific, it can be store as a document and NoSQL can handle large-scale, semi-structured data.

### (b)

NoSQL - Key-Value: Key-Value stores are in-memory databased optimized for low-latency, requires fast read and lookup.

### (c)

SQL

### (d)

NoSQL - Wide-Column: The data is typically high volume and schema flexibility helps adapt as new data fields are added.

### (e)

NoSQL - Document database: Requires handling multiple active sessions at once and requires fast writes.

## Question 5)

Quorum consistency is when there is a series of nodes in a distributed system agreeing on a read or write operation. Quorum consistency is typically useful in balancing consistency and availability by allowing tunable consistency. The effect on consistency, availability, and latency is typically due either increasing or decreasing the size of the quorum.

When deciding on suitable consistency level, it is good to consider the criticality of the data consistency as well as the latency sensitivity. We also want to make sure that read and write quorums overlap so that it reads the latest writes.
